---
title: "p8105_hw2_yh3429"
Author: "Yujie Huang"
output: github_document
date: "2022-10-05"
---


```{r, message=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

### To read and clean the NYC Transit data


```{r}
NYC_transit = read_csv( "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
   janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

```

### A short description for this dataset

#### What variables the dataset contains
The dataset has 20 variables including line, station name, station latitude, station longitude, served routes (`Route1`-`Route11`), entry, entrance type, vending, and ADA compliance. 

#### Description of the data cleaning steps so far
After importing the dataset, we cleaned the data name, specified that `Route` columns 8-11 should be character for consistency with 1-7. And then, we converted the `Route` variable from character to a logical variable.   

#### Giving the dimension (rows x columns) of the resulting dataset. Are these data tidy?
There are 1,868 rows and 20 columns in the resulting dataset. And these data is not tidy since the route number should be a variable. So, we would need to convert `Route` variables from wide to long format.


### Answer following questions

```{r}
NYC_transit %>% 
  select(station_name, line) %>% 
  distinct
```
```{r}
NYC_transit %>% 
    filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
```{r}
NYC_transit %>% 
  filter(vending == "NO") %>% 
  pull (entry) %>% 
  mean
  
```

```{r}
NYC_transit %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

NYC_transit %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

#### Descriptions of the results from above codes
According to the results, there are 465 distinct stations are there. Among them, there are 84 stations having ADA compliant.
And there is 38% station entrances/exits without vending allow entrance.In addition, there are 60 stations serving the A train and of the stations that serve the A train, there are 17 stations having ADA compliant. 




## Problem 2

### To read,clean, and organize the data from Mr. Trash Wheel sheet

```{r}
Mr_Trash_wheel = read_excel("./data/Trash Wheel Collection Data.xlsx",
                           sheet =1,
                           range = "A2:N550",
                           skip = 1,
                           col_names = TRUE,
                           col_types = NULL,
                           na = ""
                           ) %>% 
   janitor::clean_names() %>% 
   drop_na(dumpster) %>% 
   mutate(sports_balls =  as.integer(round(sports_balls, digits = 0))) %>% 
   mutate(sheet_name="Mr")
```


### To read, clean, and organize the data from Professor Trash Wheel sheet
```{r}
Pro_Trash_wheel = read_excel ("./data/Trash Wheel Collection Data.xlsx",
                           sheet =2,
                           range = "A2:M97",
                           skip = 1,
                           col_names = TRUE,
                           col_types = NULL,
                           na = ""
                           ) %>% 
   janitor::clean_names() %>% 
   drop_na(dumpster) %>% 
   mutate(sheet_name="Pro")
```

### To add an additional variable and combine two datasets
```{r}
Pro_Trash_wheel = mutate(Pro_Trash_wheel, 
                         year = as.character(year)
                      )
comb_Trash_wheel= 
bind_rows(Mr_Trash_wheel, Pro_Trash_wheel, .id="dumpster")

```

### A paragraph description for these datasets
1. In the combined trash wheel dataset, there are **641** observations of **15** variables.The key variables includes `dumpster`, `month`, `year`, `date`, `weight_tons`, `volume_cubic_yards`, `plastic_bottles`, `polystyrene`, `cigarette_butts`,`glass_bottles`,`grocery_bags`, `chip_bags`, `sports_balls`, `homes_powered`, and `sheet_name`. In addition, the variable `sheet_name` is added in order to distinguish the two datasets with the same columns.

2. In the Mr. trash wheel dataset, there are **547** observations of **15** variables.
The key variables includes `dumpster`, `month`, `year`, `date`, `weight_tons`, `volume_cubic_yards`, `plastic_bottles`, `polystyrene`, `cigarette_butts`,
`glass_bottles`, `grocery_bags`,`chip_bags`, `sports_balls`, `homes_powered`,`sheet_name`.  

3. In the Professor trash wheel dataset, there are **94** observations of **14** variables.The key variables includes `dumpster`, `month`, `year`, `date`, `weight_tons`, `volume_cubic_yards`, `plastic_bottles`, `polystyrene`, `cigarette_butts`,`glass_bottles`, `grocery_bags`,`chip_bags`, `homes_powered`,`sheet_name`.  

#### For available data, what was the total weight of trash collected by Professor Trash Wheel?

```{r}
Total_weight =  sum (Pro_Trash_wheel$weight_tons)
```
According to the sum result, the total weight of trash collected by
Professor Trash wheel is **190.12 tons**.

#### For available data, what was the total number of sports balls collected by Mr.Trash wheel in 2020?

```{r}
Total_sports_balls= Mr_Trash_wheel %>%  filter(year=="2020") %>% 
  summarise(sum(sports_balls))
```
According to the sum result in 2020, the total number of sports balls collected by
Mr.Trash wheel is **856**. 

